Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-02-08 paperIEEE.tex
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@112 db7c848b-7755-4df2-b763-8796797afa52
units & hardware
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@113 db7c848b-7755-4df2-b763-8796797afa52

\textsf{Collage} provides networking functionality of different abstraction
layers, gradually providing higher level functionality for the programmer. The
main primitives in \textsf{Collage} and their relation are shown in
\fig{fNetObject} and provide:

@@

% Time to render a 4k, 32-step MSAA image of the 1mm David model and Livre using 1 to 10 nodes with all modes (2D, 2D LB, DB, DPlex (over AA steps), tiles, chunks, pixel, subpixel) (+DFR for Livre).

We conducted performance benchmarks for the decomposition modes on a cluster
using hexacore Intel Xeon E5-2620v3 CPUs (2.4GHz), \FIXME{GPUs}, 16 GB main
memory, 4 GBit/s Ethernet and QDR Infiniband. GCC 4.8 has been used with CMake
3.7 release mode settings to compile the software stack.

@@

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eVolve}
  \caption{\label{fBenchmarkEVolve} Flower volume model at $1024^3$ unsigned byte (1GB).}
\end{figure}

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eqPly2}
  \caption{\label{fBenchmarkEqPly} David statue at 1mm resolution (988MB).}
\end{figure}

@@

The data distribution benchmarks have been performed on a cluster using
dual processor Intel Xeon X5690 CPUs (3.47GHz), 24 GB main memory, 10GBit/s
Ethernet and QDR Infiniband. Intel ICC 2017 has been used with CMake 3.2 release
mode settings to compile the software stack.

To benchmark the data distribution we used two datasets: The David statue at 2mm
resolution (as in \fig{fBenchmarkEqPly}, but in 2mm resolution) and 3D volumes
of spike frequencies of an electrical simulation of three million neurons
(\fig{fVolumes}).

@@

\begin{figure}[ht]\center
  \includegraphics[height=.44\columnwidth]{images/spikes512}\hfil
  \includegraphics[height=.44\columnwidth]{images/beechnut}
  \caption{\label{fVolumes}Volumes used for object distribution benchmarks:
    spike frequencies of a three million neuron electrical simulation at
    512x437x240 unsigned byte (51MB, left) and an MicroCT scan of a beechnut at
    1024x1024x1546 unsigned short (3GB, right) resolution}
\end{figure}

@@

The David statue at 2mm resolution is organized in a kD-tree for rendering. Each
node is a separate distributed object, which has two child node objects. A
total of $1023$ objects is distributed and synchronized. \fig{fDavidDist} shows
the distribution times for this data set. Due to limited compressibility of the
data, the results are relatively similar. Compressing the data repeatedly for
each client leads to decreased performance, since the compression overhead
cannot be amortized by the decreased transmission time. Buffering data slightly
improves performance by reducing the CPU and copy overhead. Combining
compression and buffering leads to the best performance, although only by about
10\%.

@@

The volume data sets are distributed in a single object, serializing the raw
volume buffer. The Spike volume data set has a significant compression ratio,
which is reflected by the results in \fig{fSpikesDist}. Compression for this
data is beneficial for transmitting data over a 10~Gb/s link even for a single
slave process. Buffering has little benefit since the serialization of volume
data is trivial. Buffered compression makes a significant difference, since the
compression cost can be amortized over $n$ nodes, reaching raw data transmission
rates of 3.7 GB/s with the default Snappy compressor and at best 4.4 GB/s with
ZStandard at level 1.

The distribution of the beechnut data set also behaves as expected
(\fig{fBeechnutDist}). Due to the larger object size, uncompressed transmission
is slightly faster compared to the Spike data set at 700 MB/s, and compressed
transmission does not improve the mapping performance, likely due to increased
memory pressure caused by the large data size. The comparison of the various
compression engines is consistent with the benchmarks in
\fig{fCompressorDetail}; RLE, Snappy and the LZ variants are very close to each
other, and ZSTD1 can provide better performance after four nodes due to the
better compression ratio.

@@

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/compressorDetail}
  \caption{\label{fCompressorDetail}Data compression for ply data (left, David
    statue 2mm) and raw volumes shown in \fig{fVolumes} (middle and right)}
\end{figure*}

\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/daviddist}
  \caption{\label{fDavidDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the David 2mm data set}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/spikesdist}
  \caption{\label{fSpikesDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the Spike data set in
    \fig{fVolumes}, left}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/beechnutdist}
  \caption{\label{fBeechnutDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the beechnut data set in
    \fig{fVolumes}, right}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/ifdist}
  \caption{\label{fIFDist}Object synchronization using different network
    transports}
\end{figure*}

>>>>

\textsf{Collage} provides networking functionality of different abstraction
layers, gradually providing higher level functionality for the programmer. The
main primitives in \textsf{Collage} and their relations are shown in
\fig{fNetObject} and provide:

@@

% Time to render a 4k, 32-step MSAA image of the 1\,mm David model and Livre using 1 to 10 nodes with all modes (2D, 2D LB, DB, DPlex (over AA steps), tiles, chunks, pixel, subpixel) (+DFR for Livre).

We conducted performance benchmarks for the decomposition modes on a
rendering cluster using hexacore Intel Xeon E5-2620v3 CPUs ($2.4$\,GHz), $16$\,GB DDR4 RAM per node, and nVidia GTX $970$ GPUs with $4$\,GB VRAM each. The nodes are connected via a $40$\,Gbit/s {\em Infiniband} network.
GCC 4.8 has been used with CMake 3.7 release mode settings to compile the software stack.

@@

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eVolve}
  \caption{\label{fBenchmarkEVolve} Flower volume model at $1024^3$ unsigned byte (1\,GB).}
\end{figure}

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eqPly2}
  \caption{\label{fBenchmarkEqPly} David statue at 1\,mm resolution (988\,MB).}
\end{figure}

@@

The data distribution benchmarks have been performed on a cluster using
dual processor Intel Xeon X5690 CPUs (3.47\,GHz), 24\,GB main memory, 10\,GBit/s
Ethernet and QDR Infiniband. Intel ICC 2017 has been used with CMake 3.2 release
mode settings to compile the software stack.

To benchmark the data distribution we used two datasets: The David statue at 2\,mm
resolution (as in \fig{fBenchmarkEqPly}, but in 2\,mm resolution) and 3D volumes
of spike frequencies of an electrical simulation of three million neurons
(\fig{fVolumes}).

@@

\begin{figure}[ht]\center
  \includegraphics[height=.44\columnwidth]{images/spikes512}\hfil
  \includegraphics[height=.44\columnwidth]{images/beechnut}
  \caption{\label{fVolumes}Volumes used for object distribution benchmarks:
    spike frequencies of a three million neuron electrical simulation at
    512x437x240 unsigned byte (51\,MB, left) and an MicroCT scan of a beechnut at
    1024x1024x1546 unsigned short (3\,GB, right) resolution}
\end{figure}

@@

The David statue at 2\,mm resolution is organized in a kD-tree for rendering. Each
node is a separate distributed object, which has two child node objects. A
total of $1023$ objects is distributed and synchronized. \fig{fDavidDist} shows
the distribution times for this data set. Due to limited compressibility of the
data, the results are relatively similar. Compressing the data repeatedly for
each client leads to decreased performance, since the compression overhead
cannot be amortized by the decreased transmission time. Buffering data slightly
improves performance by reducing the CPU and copy overhead. Combining
compression and buffering leads to the best performance, although only by about
10\%.

@@

The volume data sets are distributed in a single object, serializing the raw
volume buffer. The Spike volume data set has a significant compression ratio,
which is reflected by the results in \fig{fSpikesDist}. Compression for this
data is beneficial for transmitting data over a 10\,Gb/s %why was there a tilde inbetween? Should this be ~10Gb/s instead?
link even for a single
slave process. Buffering has little benefit since the serialization of volume
data is trivial. Buffered compression makes a significant difference, since the
compression cost can be amortized over $n$ nodes, reaching raw data transmission
rates of 3.7\,GB/s with the default Snappy compressor and at best 4.4\,GB/s with
ZStandard at level 1.

The distribution of the beechnut data set also behaves as expected
(\fig{fBeechnutDist}). Due to the larger object size, uncompressed transmission
is slightly faster compared to the Spike data set at 700\,MB/s, and compressed
transmission does not improve the mapping performance, likely due to increased
memory pressure caused by the large data size. The comparison of the various
compression engines is consistent with the benchmarks in
\fig{fCompressorDetail}; RLE, Snappy and the LZ variants are very close to each
other, and ZSTD1 can provide better performance after four nodes due to the
better compression ratio.

@@

\begin{figure*}[ht]\center
  \includegraphics[width=\textwidth]{images/compressorDetail}
  \caption{\label{fCompressorDetail}Data compression for ply data (left, David
    statue 2\,mm) and raw volumes shown in \fig{fVolumes} (middle and right)}
\end{figure*}

\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/daviddist}
  \caption{\label{fDavidDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the David 2\,mm data set}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/spikesdist}
  \caption{\label{fSpikesDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the Spike data set in
    \fig{fVolumes}, left}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/beechnutdist}
  \caption{\label{fBeechnutDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the beechnut data set in
    \fig{fVolumes}, right}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/ifdist}
  \caption{\label{fIFDist}Object synchronization using different network
    transports}
\end{figure*}
